Analysis: Linear Models

### 2021-12-13 code/analysis_code/analysis-linearmodels.Rmd
Analysis Complete

* The two outcomes of interest are Requested Amount (`ReqAmt`) and Obligated Amount (`OblAmt`).
* The primary predictor of interest is Incident Duration (`IncidentDuration`)
* Whichever outcome is not currently fitted to the model will *not* be considered a predictor of interest

@@

## Load Processed Data
Load the data generated from the markdowns in the `processing_code` folder.
```{r load data}
#define file path
data_location <- here::here("data", "processed_data", "analysisdata.rds")

#load data
analysis_data <- readRDS(data_location)
```

@@

## Data Overview
To better understand the data, let's use `summarytools` to better visualize the data.
```{r data overview}
skimr::skim(analysis_data)
```

@@

```{r data subsets}
#obligated amount
obl_data <- analysis_data %>%
              dplyr::select(state, IncidentYear, IncidentMonth, declarationType, incidentType, IncidentDuration, ResponseDays, IH, PA, HM,
                            Population, Counties, TotalAgencies, FEMACSAvg, OblAmt)

#requested amount
req_data <- analysis_data %>%
              dplyr::select(state, IncidentYear, IncidentMonth, declarationType, incidentType, IncidentDuration, ResponseDays, IH, PA, HM,
                            Population, Counties, TotalAgencies, FEMACSAvg, ReqAmt)
```

@@

#define recipe
req_SLR_rec <- recipes::recipe(ReqAmt ~ IncidentDuration, data = req_train_data)

@@

#using the prediction capacity
req_SLR_train %>% yardstick::metrics(truth = ReqAmt, estimate = .pred)
```
This is a terrible model. Time to try adding more predictors.

@@

#define recipe
req_MVR_rec <- recipes::recipe(ReqAmt ~ ., data = req_train_data)

@@

#using the prediction capacity
req_MVR_train %>% yardstick::metrics(truth = ReqAmt, estimate = .pred)
```
While this is an improvement, it's still not a great model. Let's compare the two side-by-side using RMSE.

@@

#Use the yardstick package to calculate the metrics on resamples
req_lm_metrics <- req_lm_preds %>%
                      dplyr::group_by(data, model) %>%
                      yardstick::metrics(truth = ReqAmt, estimate = .pred)

@@

#define recipe
obl_SLR_rec <- recipes::recipe(OblAmt ~ IncidentDuration, data = obl_train_data)

@@

#using the prediction capacity
obl_SLR_train %>% yardstick::metrics(truth = OblAmt, estimate = .pred)
```
This is a terrible model. Time to try adding more predictors.

@@

#define recipe
obl_MVR_rec <- recipes::recipe(OblAmt ~ ., data = obl_train_data)

@@

#using the prediction capacity
obl_MVR_train %>% yardstick::metrics(truth = OblAmt, estimate = .pred)
```
While this is an improvement, it's still not a great model. Let's compare the two side-by-side using RMSE.

@@

#Use the yardstick package to calculate the metrics on resamples
obl_lm_metrics <- obl_lm_preds %>%
                      dplyr::group_by(data, model) %>%
                      yardstick::metrics(truth = OblAmt, estimate = .pred)

>>>>

* The two outcomes of interest are Requested Amount (`ReqAmt`) and Obligated Amount (`OblAmt`).
* The primary predictor of interest is Incident Duration (`IncidentDuration`)
* Whichever outcome is not currently fitted to the model will *not* be considered a predictor of interest
* The outcomes of interest are log-transformed

@@

## Load Processed Data
Load the data generated from the markdowns in the `processing_code` folder.
```{r load data}
#define file path to req data
data_location_req <- here::here("data", "processed_data", "analysisdata-RA.rds")

#load req data
analysis_data_req <- readRDS(data_location_req)

#define file path to obl data
data_location_obl <- here::here("data", "processed_data", "analysisdata-OA.rds")

#load obl data
analysis_data_obl <- readRDS(data_location_obl)
```

@@

## Data Overview
To better understand the data, let's use `summarytools` to better visualize the data.
```{r data overview}
skimr::skim(analysis_data_req)

skimr::skim(analysis_data_obl)
```

@@

```{r data subsets}
#log transform ReqAmt
analysis_data_req$logReqAmt <- log(analysis_data_req$ReqAmt)

#drop unnecessary variables
req_data <- analysis_data_req %>%
              dplyr::select(-c(ReqAmt, disasterNumber, declarationTitle))

#log transform OblAmt
analysis_data_obl$logOblAmt <- log(analysis_data_obl$OblAmt)

#drop unnecessary variables
obl_data <- analysis_data_obl %>%
              dplyr::select(-c(OblAmt, disasterNumber, declarationTitle))
```

@@

#define recipe
req_SLR_rec <- recipes::recipe(logReqAmt ~ IncidentDuration, data = req_train_data)

@@

#using the prediction capacity
req_SLR_train %>% yardstick::metrics(truth = logReqAmt, estimate = .pred)
```
This isn't a great model. Time to try adding more predictors.

@@

#define recipe
req_MVR_rec <- recipes::recipe(logReqAmt ~ ., data = req_train_data)

@@

#using the prediction capacity
req_MVR_train %>% yardstick::metrics(truth = logReqAmt, estimate = .pred)
```
While this is an improvement, it's still not a great model. Let's compare the two side-by-side using RMSE.

@@

#Use the yardstick package to calculate the metrics on resamples
req_lm_metrics <- req_lm_preds %>%
                      dplyr::group_by(data, model) %>%
                      yardstick::metrics(truth = logReqAmt, estimate = .pred)

@@

#define recipe
obl_SLR_rec <- recipes::recipe(logOblAmt ~ IncidentDuration, data = obl_train_data)

@@

#using the prediction capacity
obl_SLR_train %>% yardstick::metrics(truth = logOblAmt, estimate = .pred)
```
This isn't a great model. Time to try adding more predictors.

@@

#define recipe
obl_MVR_rec <- recipes::recipe(logOblAmt ~ ., data = obl_train_data)

@@

#using the prediction capacity
obl_MVR_train %>% yardstick::metrics(truth = logOblAmt, estimate = .pred)
```
While this is an improvement, it's still not a great model. Let's compare the two side-by-side using RMSE.

@@

#Use the yardstick package to calculate the metrics on resamples
obl_lm_metrics <- obl_lm_preds %>%
                      dplyr::group_by(data, model) %>%
                      yardstick::metrics(truth = logOblAmt, estimate = .pred)
